---
title: "all_facilities"
author: "Sarah Tillman"
date: "4/27/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# library & packages
```{r load libraries, error=FALSE, message=FALSE}
library(plyr)
library(magrittr)
library(readr)
library(tidyverse)
library(gridExtra)
library(rstan)
library(bayesplot)
library(loo)
library(rstanarm)
library(bayestestR)
```
# Cases & Deaths Data
```{r load data, message=FALSE}
mask_use <- read_csv(file="../mask-use/mask-use-by-county.csv")
cases_recent <- read_csv(file="../us-states.csv")
prison_facilities <- read_csv(file="../prisons/facilities.csv")
systems <- read.csv(file="../prisons/systems.csv")
```
# State Totals at End of March
```{r}
state.cases <- cases_recent %>%
  filter(date == "2021-03-31") %>%
  select(-c("date", "fips"))
```
# Prison Systems Data
```{r join with statewide data}
# drop unimportant columns & add overall state data
systems <- systems %>%
  inner_join(state.cases,by=c("system" = "state")) %>%
  select(-c(max_inmate_population_2020))
# rename state data columns
colnames(systems)[8] <- "state_cases"
colnames(systems)[9] <- "state_deaths"
# drop NA's
systems <- drop_na(systems)
```
```{r scale variables}
# scale numerical values
systems$latest_inmate_population <- scale(systems$latest_inmate_population)
systems$total_inmate_cases <- scale(systems$total_inmate_cases)
systems$total_inmate_deaths <- scale(systems$total_inmate_deaths)
systems$total_officer_cases <- scale(systems$total_officer_cases)
systems$total_officer_deaths <- scale(systems$total_officer_deaths)
systems$inmate_tests <- scale(systems$inmate_tests)
systems$state_cases <- scale(systems$state_cases)
systems$state_deaths <- scale(systems$state_deaths)
```
```{r initial response variable graphs}
systems %>%
  ggplot(aes(x=total_inmate_deaths)) + geom_density()
p1 <- systems %>%
  ggplot(aes(x=latest_inmate_population, y = total_inmate_deaths)) + geom_point() +
  labs(x = "Latest Inmate Population", y = "Inmate Deaths")
p2 <- systems %>%
  ggplot(aes(x=total_inmate_cases, y = total_inmate_deaths)) + geom_point() +
  labs(x = "Total Inmate Cases", y = "Inmate Deaths")
p3 <- systems %>%
  ggplot(aes(x=total_officer_cases, y = total_inmate_deaths)) + geom_point() +
  labs(x = "Total Officer Cases", y = "Inmate Deaths")
p4 <- systems %>%
  ggplot(aes(x=total_officer_deaths, y = total_inmate_deaths)) + geom_point() +
  labs(x = "Total Officer Deaths", y = "Inmate Deaths")
p5 <- systems %>%
  ggplot(aes(x=inmate_tests, y = total_inmate_deaths)) + geom_point() +
  labs(x = "Inmate Tests", y = "Inmate Deaths")
p6 <- systems %>%
  ggplot(aes(x=state_deaths, y = total_inmate_deaths)) + geom_point() +
  labs(x = "State Deaths", y = "Inmate Deaths")
p7 <- systems %>%
  ggplot(aes(x=state_cases, y = total_inmate_deaths)) + geom_point() +
  labs(x = "State Cases", y = "Inmate Deaths")

grid.arrange(p1, p2, p3, p4, p5, p6, p7, ncol=2)
```
Graph appears to be right-skewed with most values around 0.
```{r flat prior}
sys.data.only <- systems %>%
  select(-c("system"))
flat.prior <- stan_glm(data = sys.data.only, formula = total_inmate_deaths~., prior=NULL,
                       family= gaussian(),
                         refresh = 0, refresh = 0, seed = 1)
posterior.flat.prior <- as.matrix(flat.prior)
mcmc_areas(posterior.flat.prior, prob=0.90)

color_scheme_set("red")
ppc_dens_overlay(y = as.vector(flat.prior$y),
                 yrep = posterior_predict(flat.prior, draws = 50))
```

```{r normal prior}
normal.prior <- stan_glm(data = sys.data.only, formula = total_inmate_deaths~., prior=normal(0,1),
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 2)
posterior.normal.prior <- as.matrix(normal.prior)
mcmc_areas(posterior.normal.prior, prob=0.90)

color_scheme_set("red")
ppc_dens_overlay(y = as.vector(normal.prior$y),
                 yrep = posterior_predict(normal.prior, draws = 50))
```

```{r full interactions model}
normal.full1 <- stan_glm(data = sys.data.only, 
                        formula = total_inmate_deaths~ inmate_tests*(total_inmate_cases + latest_inmate_population+
                                                                  total_officer_cases+ total_officer_deaths+
                                                                  state_cases+ state_deaths), 
                         prior=normal(0,1),
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 4)
posterior.normal.full1 <- as.matrix(normal.full1)
mcmc_areas(posterior.normal.full1, prob=0.90)

color_scheme_set("red")
ppc_dens_overlay(y = as.vector(normal.full1$y),
                 yrep = posterior_predict(normal.full1, draws = 50))
```
```{r}
round(posterior_interval(normal.full1, prob = 0.90), 3)
```
inmate_tests:latest_inmate_population  0.116  1.532
inmate_tests:total_officer_deaths     -0.719 -0.196
```{r}
normal.full2 <- stan_glm(data = sys.data.only, 
                        formula = total_inmate_deaths~ inmate_tests*(total_officer_deaths + 
                                                                       latest_inmate_population) +
                          total_inmate_cases*(latest_inmate_population + total_officer_cases 
                                              + total_officer_deaths + state_cases + state_deaths), 
                         prior=normal(0,1),
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 4)
posterior.normal.full2 <- as.matrix(normal.full1)
mcmc_areas(posterior.normal.full2, prob=0.90)

color_scheme_set("red")
ppc_dens_overlay(y = as.vector(normal.full2$y),
                 yrep = posterior_predict(normal.full2, draws = 50))
```
```{r}
round(posterior_interval(normal.full2, prob = 0.90), 3)
```
total_inmate_cases:latest_inmate_population  0.290  1.905
total_inmate_cases:state_cases              -1.904 -0.031
```{r}
normal.full3 <- stan_glm(data = sys.data.only, 
                        formula = total_inmate_deaths~ inmate_tests*(total_officer_deaths +
                                                                       latest_inmate_population) +
                          total_inmate_cases*(latest_inmate_population + state_cases) +
                          latest_inmate_population*(total_officer_cases + total_officer_deaths + 
                                                      state_cases + state_deaths), 
                         prior=normal(0,1),
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 4)
posterior.normal.full3 <- as.matrix(normal.full3)
mcmc_areas(posterior.normal.full3, prob=0.90)

color_scheme_set("red")
ppc_dens_overlay(y = as.vector(normal.full3$y),
                 yrep = posterior_predict(normal.full3, draws = 50))
```
```{r}
round(posterior_interval(normal.full3, prob = 0.90), 3)
```
latest_inmate_population:state_deaths         -1.137 -0.078
```{r}
normal.full4 <- stan_glm(data = sys.data.only, 
                        formula = total_inmate_deaths~ inmate_tests*(total_officer_deaths +
                                                                       latest_inmate_population) +
                          total_inmate_cases*(latest_inmate_population + state_cases) +
                          latest_inmate_population*(state_deaths) +
                          total_officer_cases*(total_officer_deaths + state_cases + state_deaths), 
                         prior=normal(0,1),
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 4)
posterior.normal.full4 <- as.matrix(normal.full3)
mcmc_areas(posterior.normal.full4, prob=0.90)

color_scheme_set("red")
ppc_dens_overlay(y = as.vector(normal.full4$y),
                 yrep = posterior_predict(normal.full4, draws = 50))
```
```{r}
round(posterior_interval(normal.full4, prob = 0.90), 3)
```
state_deaths:total_officer_cases            -1.300 -0.027
```{r}
normal.full5 <- stan_glm(data = sys.data.only, 
                        formula = total_inmate_deaths~ inmate_tests*(total_officer_deaths +
                                                                       latest_inmate_population) +
                          total_inmate_cases*(latest_inmate_population + state_cases) +
                          latest_inmate_population*(state_deaths) +
                          total_officer_cases*(state_deaths) +
                          total_officer_deaths*(state_cases + state_deaths) +
                          state_cases*(state_deaths), 
                         prior=normal(0,1),
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 4)
posterior.normal.full5 <- as.matrix(normal.full5)
mcmc_areas(posterior.normal.full5, prob=0.90)

color_scheme_set("red")
ppc_dens_overlay(y = as.vector(normal.full5$y),
                 yrep = posterior_predict(normal.full5, draws = 50))
```
```{r}
round(posterior_interval(normal.full5, prob = 0.90), 3)
```
state_cases:state_deaths                    -1.123 -0.266
```{r}
loo.ints <- loo(normal.full5, save_psis = T)
loo.main <- loo(normal.prior, save_psis = T)
loo_compare(loo.ints, loo.main)
```
```{r}
sig.model <- stan_glm(data = sys.data.only, 
                        formula = total_inmate_deaths~ inmate_tests*(total_officer_deaths +
                                                                       latest_inmate_population) +
                          total_inmate_cases*(latest_inmate_population + state_cases) +
                          latest_inmate_population*(state_deaths) +
                          state_cases*(state_deaths), 
                         prior=normal(0,1),
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 4)
posterior.sig.model <- as.matrix(sig.model)
mcmc_areas(posterior.sig.model, prob=0.90)

color_scheme_set("red")
ppc_dens_overlay(y = as.vector(sig.model$y),
                 yrep = posterior_predict(sig.model, draws = 50)) 
```
```{r}
round(posterior_interval(sig.model, prob = 0.90), 3)
```
```{r}
sig2.model <- stan_glm(data = sys.data.only, 
                        formula = total_inmate_deaths~ inmate_tests*(total_officer_deaths) +
                          latest_inmate_population*(state_deaths) +
                          state_cases*(state_deaths), 
                         prior=normal(0,1),
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 4)
posterior.sig2.model <- as.matrix(sig2.model)
mcmc_areas(posterior.sig2.model, prob=0.90)

color_scheme_set("red")
ppc_dens_overlay(y = as.vector(sig2.model$y),
                 yrep = posterior_predict(sig2.model, draws = 50))
```
```{r}
round(posterior_interval(sig2.model, prob = 0.90), 3)
```
```{r}
loo.ints <- loo(sig2.model, save_psis = T)
loo.main <- loo(normal.prior, save_psis = T)
loo_compare(loo.ints, loo.main)
```
```{r}
loo.ints
loo.main
```
Looic is much lower for model with interactions compared to model with only main effects.

```{r}
plot(loo.ints, label_points=TRUE)
plot(loo.main, label_points=TRUE)
```
3, 5, 9, 22, 28, 32, 35, 37, 38, 43
```{r}

```


