---
title: "GA Only EDA"
author: "Sarah Tillman"
date: "4/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# library & packages
```{r load libraries, error=FALSE, message=FALSE}
library(plyr)
library(magrittr)
library(readr)
library(tidyverse)
library(gridExtra)

# linear model w/ g prior function
lm.gprior<-function(y,X,g=dim(X)[1],nu0=1,s20=try(summary(lm(y~-1+X))$sigma^2,silent=TRUE),S=1000)
{
  
  n<-dim(X)[1] ; p<-dim(X)[2]
  Hg<- (g/(g+1)) * X%*%solve(t(X)%*%X)%*%t(X)
  SSRg<- t(y)%*%( diag(1,nrow=n)  - Hg ) %*%y
  
  s2<-1/rgamma(S, (nu0+n)/2, (nu0*s20+SSRg)/2 )
  
  Vb<- g*solve(t(X)%*%X)/(g+1)
  Eb<- Vb%*%t(X)%*%y
  
  E<-matrix(rnorm(S*p,0,sqrt(s2)),S,p)
  beta<-t(  t(E%*%chol(Vb)) +c(Eb))
  
  list(beta=beta,s2=s2)                                
}
```

# Cases & Deaths Data
```{r load data, message=FALSE}
mask_use <- read_csv(file="../mask-use/mask-use-by-county.csv")
cases_total <- read_csv(file="../us-counties.csv")
cases_recent <- read_csv(file="../us-counties-recent.csv")
prison_facilities <- read_csv(file="../prisons/facilities.csv")
systems <- read.csv(file="../prisons/systems.csv")
```

```{r adding facility type variables}
prison_facilities <- prison_facilities %>%
  filter(facility_type != "Reservation jail" & facility_type != "U.S. Marshalls")

facility_group <- rep(0, nrow(prison_facilities))
prison_facilities["facility_group"] <- facility_group
prison <- rep(0, nrow(prison_facilities))
prison_facilities["prison"] <- prison

prison_facilities <- prison_facilities %>%
  mutate (facility_group = case_when(
  facility_type == "Federal halfway house"~ 2,
  facility_type == "Federal prison"~ 2,
  facility_type == "State facility"~ 1,
  facility_type == "State halfway house"~ 1,
  facility_type == "State juvenile detention"~ 1,
  facility_type == "State prison"~ 1,
  facility_type == "State rehabilitation center"~ 1,
  facility_type == "State work camp"~ 1,
  TRUE ~ 0)) %>%
  mutate(prison = case_when(
    facility_type == "State prison"~ 1,
    facility_type == "Federal prison"~ 1,
    TRUE ~ 0))

group_by(prison_facilities, facility_group, facility_type) %>%
  summarise(count = n())
```

```{r dropping unimportant data & NA's}
# drop unimportant columns
ga_facilities <- prison_facilities %>%
  select(-c(nyt_id, facility_lng, facility_lat, max_inmate_population_2020, note, facility_city,
            total_officer_deaths, facility_name)) %>%
  filter(facility_state == "Georgia")

ga_facilities <- drop_na(ga_facilities)

group_by(ga_facilities, facility_type) %>%
  summarise(count = n())
```

```{r adding metro variable}
ga_facilities["metro"] <- rep(0, nrow(ga_facilities))

ga_facilities <- ga_facilities %>%
  mutate (metro = case_when(
  facility_county == "Fulton"~ 1,
  facility_county == "Cobb"~ 1,
  facility_county == "Gwinnett"~ 1,
  facility_county == "Clayton"~ 1,
  facility_county == "DeKalb"~ 1,
  facility_county == "Coweta"~ 1,
  facility_county == "Fayette"~ 1,
  facility_county == "Henry"~ 1,
  facility_county == "Douglas"~ 1,
  TRUE ~ 0))
group_by(ga_facilities, metro) %>%
  summarise(count= n())

colnames(ga_facilities)[3] <- "fips"
colnames(mask_use)[1] <- "fips"
ga_facilities <- inner_join(ga_facilities, mask_use, by = "fips")
```

```{r adding rate variables}
ga_facilities["inmate_mortality"] <- ga_facilities$total_inmate_deaths / ga_facilities$total_inmate_cases * 100
ga_facilities["inmate_infection_rate"] <- ga_facilities$total_inmate_cases / ga_facilities$latest_inmate_population * 100
ga_facilities["officer_infection_rate"] <- ga_facilities$total_officer_cases / ga_facilities$latest_inmate_population * 100

ga_facilities[is.na(ga_facilities)] <- 0
```

```{r response variables interactions}
p1 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=total_inmate_deaths, color=metro)) + geom_point() + 
  labs(x="Total Inmate Cases", y= "Total Inmate Deaths", title="Inmate Cases v. Deaths") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=total_inmate_deaths, color=prison)) + geom_point() + 
  labs(x="Total Inmate Cases", y= "Total Inmate Deaths", title="Inmate Cases v. Deaths") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=total_inmate_deaths, color=facility_group)) + geom_point() + 
  labs(x="Total Inmate Cases", y= "Total Inmate Deaths", title="Inmate Cases v. Deaths") +
  facet_grid(cols=vars(facility_group))

grid.arrange(p1, p2, p3, ncol=1)

p1 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=total_officer_cases, color=metro)) + geom_point() + 
  labs(x="Total Inmate Cases", y= "Total Officer Cases", title="Inmate Cases v. Officer Cases") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=total_officer_cases, color=prison)) + geom_point() + 
  labs(x="Total Inmate Cases", y= "Total Officer Cases", title="Inmate Cases v. OFficer Cases") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=total_officer_cases, color=facility_group)) + geom_point() + 
  labs(x="Total Inmate Cases", y= "Total Officer Cases", title="Inmate Cases v. Officer Cases") +
  facet_grid(cols=vars(facility_group))

grid.arrange(p1, p2, p3, ncol=1)

p1 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=latest_inmate_population, color=metro)) + geom_point() + 
  labs(x="Total Inmate Cases", y= "Latest Inmate Population", title="Inmate Cases v. Inmate Population") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=latest_inmate_population, color=prison)) + geom_point() + 
  labs(x="Total Inmate Cases", y= "Latest Inmate Population", title="Inmate Cases v. Inmate Population") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=latest_inmate_population, color=facility_group)) + geom_point() + 
  labs(x="Total Inmate Cases", y= "Latest Inmate Population", title="Inmate Cases v. Inmate Population") +
  facet_grid(cols=vars(facility_group))

grid.arrange(p1, p2, p3, ncol=1)

p1 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_deaths, y=latest_inmate_population, color=metro)) + geom_point() + 
  labs(x="Total Inmate Deaths", y= "Latest Inmate Population", title="Inmate Deaths v. Inmate Population") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_deaths, y=latest_inmate_population, color=prison)) + geom_point() + 
  labs(x="Total Inmate Deaths", y= "Latest Inmate Population", title="Inmate Deaths v. Inmate Population") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_deaths, y=latest_inmate_population, color=facility_group)) + geom_point() + 
  labs(x="Total Inmate Deaths", y= "Latest Inmate Population", title="Inmate Deaths v. Inmate Population") +
  facet_grid(cols=vars(facility_group))

grid.arrange(p1, p2, p3, ncol=1)

p1 <- ga_facilities %>%
  ggplot(aes(x=total_officer_cases, y=latest_inmate_population, color=metro)) + geom_point() + 
  labs(x="otal Officer Cases", y= "Latest Inmate Population", title="Officer Cases v. Inmate Population") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=total_officer_cases, y=latest_inmate_population, color=prison)) + geom_point() + 
  labs(x="otal Officer Cases", y= "Latest Inmate Population", title="Officer Cases v. Inmate Population") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=total_officer_cases, y=latest_inmate_population, color=facility_group)) + geom_point() + 
  labs(x="Total Officer Cases", y= "Latest Inmate Population", title="Officer Cases v. Inmate Population") +
  facet_grid(cols=vars(facility_group))

grid.arrange(p1, p2, p3, ncol=1)
```

```{r rate variable interactions}
p1 <- ga_facilities %>%
  ggplot(aes(x=total_officer_cases, y=inmate_mortality, color=metro)) + geom_point() + 
  labs(x="Total Officer Cases", y= "Inmate Mortality Rate", title="Officer Cases v. Inmate Mortality") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=total_officer_cases, y=inmate_mortality, color=prison)) + geom_point() + 
  labs(x="Total Officer Cases", y= "Inmate Mortality Rate", title="Officer Cases v. Inmate Mortality") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=total_officer_cases, y=inmate_mortality, color=facility_group)) + geom_point() + 
  labs(x="Total Officer Cases", y= "Inmate Mortality Rate", title="Officer Cases v. Inmate Mortality") +
  facet_grid(cols=vars(facility_group))

grid.arrange(p1, p2, p3, ncol=1)

p1 <- ga_facilities %>%
  ggplot(aes(x=latest_inmate_population, y=inmate_mortality, color=metro)) + geom_point() + 
  labs(x="Inmate Population", y= "Inmate Mortality Rate", title="Inmate Mortality v. Inmate Population") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=latest_inmate_population, y=inmate_mortality, color=prison)) + geom_point() + 
  labs(x="Inmate Population", y= "Inmate Mortality Rate", title="Inmate Mortality v. Inmate Population") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=latest_inmate_population, y=inmate_mortality, color=facility_group)) + geom_point() + 
  labs(x="Inmate Population", y= "Inmate Mortality Rate", title="Inmate Mortality v. Inmate Population") +
  facet_grid(cols=vars(facility_group))

grid.arrange(p1, p2, p3, ncol=1)

p1 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=inmate_mortality, color=metro)) + geom_point() + 
  labs(x="Inmate Cases", y= "Inmate Mortality Rate", title="Inmate Mortality v. Inmate Cases") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=inmate_mortality, color=prison)) + geom_point() + 
  labs(x="Inmate Cases", y= "Inmate Mortality Rate", title="Inmate Mortality v. Inmate Cases") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=inmate_mortality, color=facility_group)) + geom_point() + 
  labs(x="Inmate Cases", y= "Inmate Mortality Rate", title="Inmate Mortality v. Inmate Cases") +
  facet_grid(cols=vars(facility_group))

grid.arrange(p1, p2, p3, ncol=1)

p1 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=inmate_infection_rate, color=metro)) + geom_point() + 
  labs(x="Inmate Cases", y= "Inmate Infection Rate", title="Inmate Infection Rate v. Inmate Cases") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=inmate_infection_rate, color=prison)) + geom_point() + 
  labs(x="Inmate Cases", y= "Inmate Infection Rate", title="Inmate Infection Rate v. Inmate Cases") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=inmate_infection_rate, color=facility_group)) + geom_point() + 
  labs(x="Inmate Cases", y= "Inmate Infection Rate", title="Inmate Infection Rate v. Inmate Cases") +
  facet_grid(cols=vars(facility_group))

grid.arrange(p1, p2, p3, ncol=1)

p1 <- ga_facilities %>%
  ggplot(aes(x=total_officer_cases, y=inmate_infection_rate, color=metro)) + geom_point() + 
  labs(x="Officer Cases", y= "Inmate Infection Rate", title="Inmate Infection Rate v. IOfficer Cases") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=total_officer_cases, y=inmate_infection_rate, color=prison)) + geom_point() + 
  labs(x="Officer Cases", y= "Inmate Infection Rate", title="Inmate Infection Rate v. Officer Cases") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=total_officer_cases, y=inmate_infection_rate, color=facility_group)) + geom_point() + 
  labs(x="Officer Cases", y= "Inmate Infection Rate", title="Inmate Infection Rate v. Officer Cases") +
  facet_grid(cols=vars(facility_group))

grid.arrange(p1, p2, p3, ncol=1)

p1 <- ga_facilities %>%
  ggplot(aes(x=total_officer_cases, y=officer_infection_rate, color=metro)) + geom_point() + 
  labs(x="Officer Cases", y= "Officer Infection Rate", title="Officer Infection Rate v. Officer Cases") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=total_officer_cases, y=officer_infection_rate, color=prison)) + geom_point() + 
  labs(x="Officer Cases", y= "Officer Infection Rate", title="Officer Infection Rate v. Officer Cases") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=total_officer_cases, y=officer_infection_rate, color=facility_group)) + geom_point() + 
  labs(x="Officer Cases", y= "Officer Infection Rate", title="Officer Infection Rate v. Officer Cases") +
  facet_grid(cols=vars(facility_group))

grid.arrange(p1, p2, p3, ncol=1)

p1 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=officer_infection_rate, color=metro)) + geom_point() + 
  labs(x="Inmate Cases", y= "Officer Infection Rate", title="Officer Infection Rate v. Inmate Cases") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=officer_infection_rate, color=prison)) + geom_point() + 
  labs(x="Inmate Cases", y= "Officer Infection Rate", title="Officer Infection Rate v. Inmate Cases") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=total_inmate_cases, y=officer_infection_rate, color=facility_group)) + geom_point() + 
  labs(x="Inmate Cases", y= "Officer Infection Rate", title="Officer Infection Rate v. Inmate Cases") +
  facet_grid(cols=vars(facility_group))

grid.arrange(p1, p2, p3, ncol=1)
```

```{r local mask usage & facility infection/mortality rates}
p1 <- ga_facilities %>%
  ggplot(aes(x=NEVER, y=officer_infection_rate, color=metro)) + geom_point() + 
  labs(x="NEVER", y= "Officer Infection Rate", title="NEVER v. Officer Infection Rate") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=NEVER, y=officer_infection_rate, color=prison)) + geom_point() + 
  labs(x="NEVER", y= "Officer Infection Rate", title="NEVER v. Officer Infection Rate") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=NEVER, y=inmate_infection_rate, color=metro)) + geom_point() + 
  labs(x="NEVER", y= "Inmate Infection Rate", title="NEVER v. Inmate Infection Rate") +
  facet_grid(cols=vars(metro))
p4 <- ga_facilities %>%
  ggplot(aes(x=NEVER, y=inmate_infection_rate, color=prison)) + geom_point() + 
  labs(x="NEVER", y= "Inmate Infection Rate", title="NEVER v. Inmate Infection Rate") +
  facet_grid(cols=vars(prison))
p5 <- ga_facilities %>%
  ggplot(aes(x=NEVER, y=inmate_mortality, color=metro)) + geom_point() + 
  labs(x="NEVER", y= "Inmate Mortality Rate", title="NEVER v. Inmate Mortality Rate") +
  facet_grid(cols=vars(metro))
p6 <- ga_facilities %>%
  ggplot(aes(x=NEVER, y=inmate_mortality, color=prison)) + geom_point() + 
  labs(x="NEVER", y= "Inmate Mortality Rate", title="NEVER v. Inmate Mortality Rate") +
  facet_grid(cols=vars(prison))

grid.arrange(p1, p2, p3, p4, p5, p6, ncol=2)

p1 <- ga_facilities %>%
  ggplot(aes(x=RARELY, y=officer_infection_rate, color=metro)) + geom_point() + 
  labs(x="RARELY", y= "Officer Infection Rate", title="RARELY v. Officer Infection Rate") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=RARELY, y=officer_infection_rate, color=prison)) + geom_point() + 
  labs(x="RARELY", y= "Officer Infection Rate", title="RARELY v. Officer Infection Rate") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=RARELY, y=inmate_infection_rate, color=metro)) + geom_point() + 
  labs(x="RARELY", y= "Inmate Infection Rate", title="RARELY v. Inmate Infection Rate") +
  facet_grid(cols=vars(metro))
p4 <- ga_facilities %>%
  ggplot(aes(x=RARELY, y=inmate_infection_rate, color=prison)) + geom_point() + 
  labs(x="RARELY", y= "Inmate Infection Rate", title="RARELY v. Inmate Infection Rate") +
  facet_grid(cols=vars(prison))
p5 <- ga_facilities %>%
  ggplot(aes(x=RARELY, y=inmate_mortality, color=metro)) + geom_point() + 
  labs(x="RARELY", y= "Inmate Mortality Rate", title="RARELY v. Inmate Mortality Rate") +
  facet_grid(cols=vars(metro))
p6 <- ga_facilities %>%
  ggplot(aes(x=RARELY, y=inmate_mortality, color=prison)) + geom_point() + 
  labs(x="RARELY", y= "Inmate Mortality Rate", title="RARELY v. Inmate Mortality Rate") +
  facet_grid(cols=vars(prison))

grid.arrange(p1, p2, p3, p4, p5, p6, ncol=2)

p1 <- ga_facilities %>%
  ggplot(aes(x=SOMETIMES, y=officer_infection_rate, color=metro)) + geom_point() + 
  labs(x="SOMETIMES", y= "Officer Infection Rate", title="SOMETIMES v. Officer Infection Rate") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=SOMETIMES, y=officer_infection_rate, color=prison)) + geom_point() + 
  labs(x="SOMETIMES", y= "Officer Infection Rate", title="SOMETIMES v. Officer Infection Rate") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=SOMETIMES, y=inmate_infection_rate, color=metro)) + geom_point() + 
  labs(x="SOMETIMES", y= "Inmate Infection Rate", title="SOMETIMES v. Inmate Infection Rate") +
  facet_grid(cols=vars(metro))
p4 <- ga_facilities %>%
  ggplot(aes(x=SOMETIMES, y=inmate_infection_rate, color=prison)) + geom_point() + 
  labs(x="SOMETIMES", y= "Inmate Infection Rate", title="SOMETIMES v. Inmate Infection Rate") +
  facet_grid(cols=vars(prison))
p5 <- ga_facilities %>%
  ggplot(aes(x=SOMETIMES, y=inmate_mortality, color=metro)) + geom_point() + 
  labs(x="SOMETIMES", y= "Inmate Mortality Rate", title="SOMETIMES v. Inmate Mortality Rate") +
  facet_grid(cols=vars(metro))
p6 <- ga_facilities %>%
  ggplot(aes(x=SOMETIMES, y=inmate_mortality, color=prison)) + geom_point() + 
  labs(x="SOMETIMES", y= "Inmate Mortality Rate", title="SOMETIMES v. Inmate Mortality Rate") +
  facet_grid(cols=vars(prison))

grid.arrange(p1, p2, p3, p4, p5, p6, ncol=2)

p1 <- ga_facilities %>%
  ggplot(aes(x=FREQUENTLY, y=officer_infection_rate, color=metro)) + geom_point() + 
  labs(x="FREQUENTLY", y= "Officer Infection Rate", title="FREQUENTLY v. Officer Infection Rate") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=FREQUENTLY, y=officer_infection_rate, color=prison)) + geom_point() + 
  labs(x="FREQUENTLY", y= "Officer Infection Rate", title="FREQUENTLY v. Officer Infection Rate") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=FREQUENTLY, y=inmate_infection_rate, color=metro)) + geom_point() + 
  labs(x="FREQUENTLY", y= "Inmate Infection Rate", title="FREQUENTLY v. Inmate Infection Rate") +
  facet_grid(cols=vars(metro))
p4 <- ga_facilities %>%
  ggplot(aes(x=FREQUENTLY, y=inmate_infection_rate, color=prison)) + geom_point() + 
  labs(x="FREQUENTLY", y= "Inmate Infection Rate", title="FREQUENTLY v. Inmate Infection Rate") +
  facet_grid(cols=vars(prison))
p5 <- ga_facilities %>%
  ggplot(aes(x=FREQUENTLY, y=inmate_mortality, color=metro)) + geom_point() + 
  labs(x="FREQUENTLY", y= "Inmate Mortality Rate", title="FREQUENTLY v. Inmate Mortality Rate") +
  facet_grid(cols=vars(metro))
p6 <- ga_facilities %>%
  ggplot(aes(x=FREQUENTLY, y=inmate_mortality, color=prison)) + geom_point() + 
  labs(x="FREQUENTLY", y= "Inmate Mortality Rate", title="FREQUENTLY v. Inmate Mortality Rate") +
  facet_grid(cols=vars(prison))

grid.arrange(p1, p2, p3, p4, p5, p6, ncol=2)

p1 <- ga_facilities %>%
  ggplot(aes(x=ALWAYS, y=officer_infection_rate, color=metro)) + geom_point() + 
  labs(x="ALWAYS", y= "Officer Infection Rate", title="ALWAYS v. Officer Infection Rate") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=ALWAYS, y=officer_infection_rate, color=prison)) + geom_point() + 
  labs(x="ALWAYS", y= "Officer Infection Rate", title="ALWAYS v. Officer Infection Rate") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=ALWAYS, y=inmate_infection_rate, color=metro)) + geom_point() + 
  labs(x="ALWAYS", y= "Inmate Infection Rate", title="ALWAYS v. Inmate Infection Rate") +
  facet_grid(cols=vars(metro))
p4 <- ga_facilities %>%
  ggplot(aes(x=ALWAYS, y=inmate_infection_rate, color=prison)) + geom_point() + 
  labs(x="ALWAYS", y= "Inmate Infection Rate", title="ALWAYS v. Inmate Infection Rate") +
  facet_grid(cols=vars(prison))
p5 <- ga_facilities %>%
  ggplot(aes(x=ALWAYS, y=inmate_mortality, color=metro)) + geom_point() + 
  labs(x="ALWAYS", y= "Inmate Mortality Rate", title="ALWAYS v. Inmate Mortality Rate") +
  facet_grid(cols=vars(metro))
p6 <- ga_facilities %>%
  ggplot(aes(x=ALWAYS, y=inmate_mortality, color=prison)) + geom_point() + 
  labs(x="ALWAYS", y= "Inmate Mortality Rate", title="ALWAYS v. Inmate Mortality Rate") +
  facet_grid(cols=vars(prison))

grid.arrange(p1, p2, p3, p4, p5, p6, ncol=2)

p1 <- ga_facilities %>%
  ggplot(aes(x=NEVER+RARELY, y=officer_infection_rate, color=metro)) + geom_point() + 
  labs(x="NEVER+RARELY", y= "Officer Infection Rate", title="NEVER+RARELY v. Officer Infection Rate") +
  facet_grid(cols=vars(metro))
p2 <- ga_facilities %>%
  ggplot(aes(x=NEVER+RARELY, y=officer_infection_rate, color=prison)) + geom_point() + 
  labs(x="NEVER+RARELY", y= "Officer Infection Rate", title="NEVER+RARELY v. Officer Infection Rate") +
  facet_grid(cols=vars(prison))
p3 <- ga_facilities %>%
  ggplot(aes(x=NEVER+RARELY, y=inmate_infection_rate, color=metro)) + geom_point() + 
  labs(x="NEVER+RARELY", y= "Inmate Infection Rate", title="NEVER+RARELY v. Inmate Infection Rate") +
  facet_grid(cols=vars(metro))
p4 <- ga_facilities %>%
  ggplot(aes(x=NEVER+RARELY, y=inmate_infection_rate, color=prison)) + geom_point() + 
  labs(x="NEVER+RARELY", y= "Inmate Infection Rate", title="NEVER+RARELY v. Inmate Infection Rate") +
  facet_grid(cols=vars(prison))
p5 <- ga_facilities %>%
  ggplot(aes(x=NEVER+RARELY, y=inmate_mortality, color=metro)) + geom_point() + 
  labs(x="NEVER+RARELY", y= "Inmate Mortality Rate", title="NEVER+RARELY v. Inmate Mortality Rate") +
  facet_grid(cols=vars(metro))
p6 <- ga_facilities %>%
  ggplot(aes(x=RARELY, y=inmate_mortality, color=prison)) + geom_point() + 
  labs(x="NEVER+RARELY", y= "Inmate Mortality Rate", title="NEVER+RARELY v. Inmate Mortality Rate") +
  facet_grid(cols=vars(prison))

grid.arrange(p1, p2, p3, p4, p5, p6, ncol=2)
```

# Regression Modeling
```{r set up full models}
# inmate mortality model
y.im <- inmate.mortality <- ga_facilities$inmate_mortality
X.im.full <- select(ga_facilities, -c("facility_type", "facility_county", "fips", "facility_state", "facility_group",
                                      "inmate_mortality", "total_inmate_deaths"))
# inmate infection rate model
y.iir <- inmate.infections <- ga_facilities$inmate_infection_rate
X.iir.full <- select(ga_facilities, -c("facility_type", "facility_county", "fips", "facility_state", "facility_group",
                                       "inmate_infection_rate", "total_inmate_deaths"))
```
```{r inmate mortality full linear models w g prior}
# inmate mortality model w/ g prior
X.im.full["intercept"] <- rep(1, nrow(X.im.full))
p <- ncol(X.im.full)
X <- as.matrix(X.im.full)
im.model.full <-lm.gprior(y.im,X,g=dim(X)[1],nu0=1,S=10000)
Beta.post <- im.model.full$beta
# Posterior mean + 90% Credible Interval
post.data <- matrix(1:(p*3), nrow=p, dimnames = list(colnames(X.im.full), c("Posterior.Mean", "CI.Lower", "CI.Upper"))) 
for(i in 1:p){
  post.mean <- mean(Beta.post[,i])
  post.interval <- quantile(Beta.post[,i], c(0.05, 0.95))
  post.data[i, ] <- c(post.mean, post.interval[1], post.interval[2])
}

data.frame(post.data)
```
latest_inmate_population --> +
total_inmate_cases --> -
prison --> +
metro --> -
```{r inmate infection rate full linear model w g prior}
# inmate infection rate model w/ g prior
X.iir.full["intercept"] <- rep(1, nrow(X.iir.full))
p <- ncol(X.iir.full)
X <- as.matrix(X.iir.full)
iir.model.full <-lm.gprior(y.iir,X,g=dim(X)[1],nu0=1,S=10000)
Beta.post <- iir.model.full$beta
# Posterior mean + 90% Credible Interval
post.data <- matrix(1:(p*3), nrow=p, dimnames = list(colnames(X.iir.full), c("Posterior.Mean", "CI.Lower", "CI.Upper"))) 
for(i in 1:p){
  post.mean <- mean(Beta.post[,i])
  post.interval <- quantile(Beta.post[,i], c(0.05, 0.95))
  post.data[i, ] <- c(post.mean, post.interval[1], post.interval[2])
}

data.frame(post.data)
```
inmate_infection_rate --> -
total_inmate_cases --> +
metro --> +
officer_infection_rate --> +
```{r inmate mortality model w mask usage groups w lm.gprior}
# inmate mortality model w/ g prior
X.im.full["NEVER+"] <- X.im.full$NEVER + X.im.full$RARELY + X.im.full$SOMETIMES
X.im.mask.grouped <- select(X.im.full, -c("NEVER", "RARELY", "SOMETIMES", "FREQUENTLY"))
p <- ncol(X.im.mask.grouped)
X <- as.matrix(X.im.mask.grouped)
im.model.full <-lm.gprior(y.im,X,g=dim(X)[1],nu0=1,S=10000)
Beta.post <- im.model.full$beta
# Posterior mean + 90% Credible Interval
post.data <- matrix(1:(p*3), nrow=p, dimnames = list(colnames(X.im.mask.grouped), c("Posterior.Mean", "CI.Lower", "CI.Upper"))) 
for(i in 1:p){
  post.mean <- mean(Beta.post[,i])
  post.interval <- quantile(Beta.post[,i], c(0.05, 0.95))
  post.data[i, ] <- c(post.mean, post.interval[1], post.interval[2])
}

data.frame(post.data)
```
Same basic results as w individual mask usage terms, but with smaller variance
```{r inmate infection rate model w mask usage groups w lm.gprior}
# inmate mortality model w/ g prior
X.iir.full["NEVER+"] <- X.iir.full$NEVER + X.iir.full$RARELY + X.iir.full$SOMETIMES
X.iir.mask.grouped <- select(X.iir.full, -c("NEVER", "RARELY", "SOMETIMES", "FREQUENTLY"))
p <- ncol(X.iir.mask.grouped)
X <- as.matrix(X.iir.mask.grouped)
iir.model.full <-lm.gprior(y.iir,X,g=dim(X)[1],nu0=1,S=10000)
Beta.post <- iir.model.full$beta
# Posterior mean + 90% Credible Interval
post.data <- matrix(1:(p*3), nrow=p, dimnames = list(colnames(X.iir.mask.grouped), c("Posterior.Mean", "CI.Lower", "CI.Upper"))) 
for(i in 1:p){
  post.mean <- mean(Beta.post[,i])
  post.interval <- quantile(Beta.post[,i], c(0.05, 0.95))
  post.data[i, ] <- c(post.mean, post.interval[1], post.interval[2])
}

data.frame(post.data)
```
Same basic results as w individual mask usage terms, but with smaller variance
```{r inmate mortality interaction terms model w lm.gprior}
# inmate mortality model w/ g prior
X.im.interactions <- X.im.mask.grouped
X.im.interactions["inmate.pop.X.prison"] <- X.im.mask.grouped$latest_inmate_population * X.im.mask.grouped$prison
X.im.interactions["inmate.pop.X.metro"] <- X.im.mask.grouped$latest_inmate_population * X.im.mask.grouped$metro
X.im.interactions["inmate.pop.X.NEVER"] <- X.im.mask.grouped$latest_inmate_population * X.im.mask.grouped$`NEVER+`
X.im.interactions["inmate.pop.X.ALWAYS"] <- X.im.mask.grouped$latest_inmate_population * X.im.mask.grouped$ALWAYS
X.im.interactions["inmate.cases.X.prison"] <- X.im.mask.grouped$total_inmate_cases * X.im.mask.grouped$prison
X.im.interactions["inmate.cases.X.metro"] <- X.im.mask.grouped$total_inmate_cases * X.im.mask.grouped$metro
X.im.interactions["inmate.cases.X.NEVER"] <- X.im.mask.grouped$total_inmate_cases * X.im.mask.grouped$`NEVER+`
X.im.interactions["inmate.cases.X.ALWAYS"] <- X.im.mask.grouped$total_inmate_cases * X.im.mask.grouped$ALWAYS
X.im.interactions["inmate.cases.X.officer.cases"] <- X.im.mask.grouped$total_inmate_cases *
  X.im.mask.grouped$total_officer_cases
X.im.interactions["officer.cases.X.prison"] <- X.im.mask.grouped$total_officer_cases * X.im.mask.grouped$prison
X.im.interactions["officer.cases.X.metro"] <- X.im.mask.grouped$total_officer_cases * X.im.mask.grouped$metro
X.im.interactions["officer.cases.X.NEVER"] <- X.im.mask.grouped$total_officer_cases * X.im.mask.grouped$`NEVER+`
X.im.interactions["officer.cases.X.ALWAYS"] <- X.im.mask.grouped$total_officer_cases * X.im.mask.grouped$ALWAYS

p <- ncol(X.im.interactions)
X <- as.matrix(X.im.interactions)
im.model.int <-lm.gprior(y.im,X,g=dim(X)[1],nu0=1,S=10000)
Beta.post <- im.model.int$beta
# Posterior mean + 90% Credible Interval
post.data <- matrix(1:(p*3), nrow=p, dimnames = list(colnames(X.im.interactions), c("Posterior.Mean", "CI.Lower", "CI.Upper"))) 
for(i in 1:p){
  post.mean <- mean(Beta.post[,i])
  post.interval <- quantile(Beta.post[,i], c(0.05, 0.95))
  post.data[i, ] <- c(post.mean, post.interval[1], post.interval[2])
}

data.frame(post.data)
```
latest_inmate_population --> +
total_inmate_cases --> -
total_officer_cases --> -
inmate_population * NEVER --> -
inmate_population * ALWAYS --> -
inmate_cases * prison --> -
inmate_cases * NEVER --> +
inmate_cases * ALWAYS --> +
officer_cases * prison --> -
officer_cases * metro --> -
officer_cases * NEVER --> +
officer_cases * ALWAYS --> +
```{r inmate infection rate interaction terms model w lm.gprior}
X.iir.interactions <- X.iir.mask.grouped
X.iir.interactions["inmate.pop.X.prison"] <- X.iir.mask.grouped$latest_inmate_population * X.iir.mask.grouped$prison
X.iir.interactions["inmate.pop.X.metro"] <- X.iir.mask.grouped$latest_inmate_population * X.iir.mask.grouped$metro
X.iir.interactions["inmate.pop.X.NEVER"] <- X.iir.mask.grouped$latest_inmate_population * X.iir.mask.grouped$`NEVER+`
X.iir.interactions["inmate.pop.X.ALWAYS"] <- X.iir.mask.grouped$latest_inmate_population * X.iir.mask.grouped$ALWAYS
X.iir.interactions["inmate.cases.X.prison"] <- X.iir.mask.grouped$total_inmate_cases * X.iir.mask.grouped$prison
X.iir.interactions["inmate.cases.X.metro"] <- X.iir.mask.grouped$total_inmate_cases * X.iir.mask.grouped$metro
X.iir.interactions["inmate.cases.X.NEVER"] <- X.iir.mask.grouped$total_inmate_cases * X.iir.mask.grouped$`NEVER+`
X.iir.interactions["inmate.cases.X.ALWAYS"] <- X.iir.mask.grouped$total_inmate_cases * X.iir.mask.grouped$ALWAYS
X.iir.interactions["inmate.cases.X.officer.cases"] <- X.iir.mask.grouped$total_inmate_cases *
  X.iir.mask.grouped$total_officer_cases
X.iir.interactions["officer.cases.X.prison"] <- X.iir.mask.grouped$total_officer_cases * X.iir.mask.grouped$prison
X.iir.interactions["officer.cases.X.metro"] <- X.iir.mask.grouped$total_officer_cases * X.iir.mask.grouped$metro
X.iir.interactions["officer.cases.X.NEVER"] <- X.iir.mask.grouped$total_officer_cases * X.iir.mask.grouped$`NEVER+`
X.iir.interactions["officer.cases.X.ALWAYS"] <- X.iir.mask.grouped$total_officer_cases * X.iir.mask.grouped$ALWAYS
X.iir.interactions["officer.ir.X.prison"] <- X.iir.mask.grouped$officer_infection_rate * X.iir.mask.grouped$prison
X.iir.interactions["officer.ir.X.metro"] <- X.iir.mask.grouped$officer_infection_rate * X.iir.mask.grouped$metro
X.iir.interactions["officer.ir.X.NEVER"] <- X.iir.mask.grouped$officer_infection_rate * X.iir.mask.grouped$`NEVER+`
X.iir.interactions["officer.ir.X.ALWAYS"] <- X.iir.mask.grouped$officer_infection_rate * X.iir.mask.grouped$ALWAYS
X.iir.interactions["officer.ir.X.inmate.cases"] <- X.iir.mask.grouped$officer_infection_rate *
  X.iir.mask.grouped$total_inmate_cases



p <- ncol(X.iir.interactions)
X <- as.matrix(X.iir.interactions)
iir.model.int <-lm.gprior(y.iir,X,g=dim(X)[1],nu0=1,S=10000)
Beta.post <- iir.model.int$beta
# Posterior mean + 90% Credible Interval
post.data <- matrix(1:(p*3), nrow=p, dimnames = list(colnames(X.iir.interactions), c("Posterior.Mean", "CI.Lower", "CI.Upper"))) 
for(i in 1:p){
  post.mean <- mean(Beta.post[,i])
  post.interval <- quantile(Beta.post[,i], c(0.05, 0.95))
  post.data[i, ] <- c(post.mean, post.interval[1], post.interval[2])
}

data.frame(post.data)
```
latest_inmate_population --> -
total_inmate_cases --> +
metro --> +
ALWAYS --> -
intercept --> +
inmate_population * metro --> -
inmate_population * NEVER --> +
inmate_population * ALWAYS --> +
inmate_cases * metro --> +
inmate_cases * ALWAYS --> -
officer_cases * NEVER --> -
officer_cases * ALWAYS --> +
```{r inmate deaths full linear model with g prior}
# inmate deaths model
y.id <- ga_facilities$total_inmate_deaths
X.id.full <- select(ga_facilities, -c("facility_type", "facility_county", "fips", "facility_state", "facility_group",
                                      "inmate_mortality", "total_inmate_deaths"))
# inmate mortality model w/ g prior
X.id.full["intercept"] <- rep(1, nrow(X.id.full))
p <- ncol(X.id.full)
X <- as.matrix(X.id.full)
id.model.full <-lm.gprior(y.id,X,g=dim(X)[1],nu0=1,S=10000)
Beta.post <- id.model.full$beta
# Posterior mean + 90% Credible Interval
post.data <- matrix(1:(p*3), nrow=p, dimnames = list(colnames(X.id.full), c("Posterior.Mean", "CI.Lower", "CI.Upper"))) 
for(i in 1:p){
  post.mean <- mean(Beta.post[,i])
  post.interval <- quantile(Beta.post[,i], c(0.05, 0.95))
  post.data[i, ] <- c(post.mean, post.interval[1], post.interval[2])
}

data.frame(post.data)
```
latest_inmate_population --> +
prison --> +
metro --> -
```{r r inmate deaths w grouped mask usage lm.gprior}
X.id.full["NEVER+"] <- X.id.full$NEVER + X.id.full$RARELY + X.id.full$SOMETIMES
X.id.mask.grouped <- select(X.id.full, -c("NEVER", "RARELY", "SOMETIMES", "FREQUENTLY"))
p <- ncol(X.id.mask.grouped)
X <- as.matrix(X.id.mask.grouped)
id.mask.grouped <-lm.gprior(y.id,X,g=dim(X)[1],nu0=1,S=10000)
Beta.post <- id.mask.grouped$beta
# Posterior mean + 90% Credible Interval
post.data <- matrix(1:(p*3), nrow=p, dimnames = list(colnames(X.id.mask.grouped), c("Posterior.Mean", "CI.Lower", "CI.Upper"))) 
for(i in 1:p){
  post.mean <- mean(Beta.post[,i])
  post.interval <- quantile(Beta.post[,i], c(0.05, 0.95))
  post.data[i, ] <- c(post.mean, post.interval[1], post.interval[2])
}

data.frame(post.data)
```
Same as non-grouped model

