---
title: "Modeling COVID-19 Deaths in US Prisons"
author: "Sarah Tillman"
date: "4/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Libraries & Dataset Imports
```{r load libraries}
library(plyr)
library(magrittr)
library(readr)
library(tidyverse)
library(gridExtra)
library(rstan)
library(bayesplot)
library(loo)
library(rstanarm)
library(bayestestR)
library(brms)
```
```{r load datasets, message=FALSE}
cases_recent <- read_csv(file="../us-states.csv")
systems.full <- read.csv(file="../prisons/systems.csv")
```
# Combine Datasets & Drop Irrelevant Data
```{r filter data & join tables}
# filter for cummulative counts on 3/31/2021
state.cases <- cases_recent %>%
  filter(date == "2021-03-31") %>%
  select(-c("date", "fips"))

# drop unimportant columns & add overall state data
systems <- systems.full %>%
  inner_join(state.cases,by=c("system" = "state")) %>%
  select(-c(max_inmate_population_2020))
# rename state data columns
colnames(systems)[8] <- "state_cases"
colnames(systems)[9] <- "state_deaths"
# drop NA's
systems <- drop_na(systems)
```
# Density Distribution of total_inmate_deaths
```{r density plot of inmate death counts}
systems %>%
  ggplot(aes(x=total_inmate_deaths)) + geom_density()
```
```{r check summary statistics of response var}
systems %>%
  summarise(mean = mean(total_inmate_deaths),
            sd = sd(total_inmate_deaths),
            variance = var(total_inmate_deaths),
            IQR = IQR(total_inmate_deaths),
            outlier_limit = mean(total_inmate_deaths) + 1.5*IQR(total_inmate_deaths))
nonoutliers.systems <- systems %>%
  filter(total_inmate_deaths <= 110.4)
```
total_inmate_deaths is a count variable with a much greater variance in comparison to the mean.
# Identifying Outliers in total_inmate_deaths 
```{r create scaled dataset}
# save scaled data as sep. dataset
systems.scaled.outliers <- systems
systems.scaled.outliers$latest_inmate_population <- scale(systems$latest_inmate_population)
systems.scaled.outliers$total_inmate_cases <- scale(systems$total_inmate_cases)
systems.scaled.outliers$total_officer_cases <- scale(systems$total_officer_cases)
systems.scaled.outliers$total_officer_deaths <- scale(systems$total_officer_deaths)
systems.scaled.outliers$inmate_tests <- scale(systems$inmate_tests)
systems.scaled.outliers$state_cases <- scale(systems$state_cases)
systems.scaled.outliers$state_deaths <- scale(systems$state_deaths)
nrv.scaled <- systems.scaled.outliers
systems.scaled.outliers$total_inmate_deaths <- scale(systems$total_inmate_deaths)
```
```{r covariates only datasets} 
# save data only models for ease of formula 
sys.data <- select(systems, -c("system")) 
nonoutliers.sys.data <- select(nonoutliers.systems, -c("system")) 
sys.scaled.data <- select(systems.scaled.outliers, -c("system"))
nrv.scaled.data <- select(nrv.scaled, -c("system"))
```
# Fit Negative Bionmial & Poisson Regression Model with Poisson GLM under full dataset
```{r STAN_GLM + Bayes Factor for outlier}
nb.1 <- stan_glm(data = sys.data, formula = total_inmate_deaths~.,
                        family=neg_binomial_2(link="log"),
                        refresh = 0, refresh = 0)
(loo.nb1 <- loo(nb.1, save_psis = T))
pois.1 <- stan_glm(data = sys.data, formula = total_inmate_deaths~.,
                        family=poisson(link="log"),
                        refresh = 0, refresh = 0)
(loo.pois1 <- loo(pois.1, save_psis = T))
loo_compare(loo.pois1, loo.nb1)
```
```{r STAN_GLM + Bayes Factor for dataset with only covariates scaled}
norm.0 <- stan_glm(data = nrv.scaled.data, formula = total_inmate_deaths~.,
                        family=gaussian(),
                        refresh = 0, refresh = 0)
(loo.norm.0 <- loo(norm.0, save_psis = T))
loo.nb1
loo_compare(loo.norm.0, loo.nb1)
```
# Fit Regression Model with Gaussian Model under Scaled Dataset
```{r STAN_GLM + Bayes Factor for Scaled Normal Model}
norm.1 <- stan_glm(data = sys.scaled.data, formula = total_inmate_deaths~.,
                        family=gaussian(),
                        refresh = 0, refresh = 0)
(loo.norm1 <- loo(norm.1, save_psis = T))
loo.nb1
loo_compare(loo.norm1, loo.nb1)
```
# Comparing Model Fits
```{r}
(BF.pois <- bayesfactor_parameters(pois.1, null = 0))
(BF.nb <- bayesfactor_parameters(nb.1, null = 0))
(BF.norm <- bayesfactor_parameters(norm.1, null = 0))
```
According to the looic comparisons above, the Normal model with scaled outliers has the best fit since it has a significantly lower looic compared to both the Negative Binomial and Poisson GLMs. 

# Fit Best Model Type With All Possible Interactions
```{r interactions w/ inmate tests}
int.1 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ inmate_tests*., 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 4)
round(posterior_interval(int.1, prob = 0.90), 3)
```
```{r interactions w/ total inmate cases}
int.2 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ total_inmate_cases*., 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(int.2, prob = 0.90), 3)
```
```{r interactions w/ inmate population}
int.3 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ latest_inmate_population*., 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(int.3, prob = 0.90), 3)
```
```{r interactions w/ officer cases}
int.4 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ total_officer_cases*., 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(int.4, prob = 0.90), 3)
```
```{r interactions w/ officer deaths}
int.5 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ total_officer_deaths*., 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(int.5, prob = 0.90), 3)
```
```{r interactions w/ statewide cases}
int.6 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ state_cases*., 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(int.6, prob = 0.90), 3)
```
```{r interactions w/ statewide deaths}
int.7 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ state_deaths*., 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(int.7, prob = 0.90), 3)
```
```{r model w/ all significant interaction terms added}
full.int <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ 
                       latest_inmate_population*(inmate_tests + total_officer_deaths + state_cases + 
                                                   state_deaths + total_officer_cases) +
                       total_officer_deaths*(total_inmate_cases + total_officer_cases) + 
                       state_cases*(state_deaths), 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(full.int, prob = 0.90), 3)
(BF.full.int <- bayesfactor_parameters(full.int, null = 0))
(loo.full.int <- loo(full.int, save_psis = T))
```
```{r optimization 1}
sig1 <- stan_glm(data = sys.scaled.data, 
                        formula = 
                    total_inmate_deaths~ inmate_tests*(latest_inmate_population + total_officer_deaths) +
                    total_inmate_cases*(total_officer_deaths) +
                    latest_inmate_population*(total_officer_deaths + state_cases + state_deaths) +
                    total_officer_cases*(total_officer_deaths) +
                    state_cases*(total_officer_deaths + state_deaths), 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
(BF.sig1 <- bayesfactor_parameters(sig1, null = 0))
(loo.sig1 <- loo(sig1, save_psis = T))
loo.full.int
```
```{r optimization 2}
loo.best <- loo.sig1
sig2 <- stan_glm(data = sys.scaled.data, 
                        formula = 
                    total_inmate_deaths~ inmate_tests*(total_officer_deaths) +
                    total_inmate_cases*(total_officer_deaths) +
                    latest_inmate_population*(total_officer_deaths + state_cases + state_deaths) +
                    total_officer_cases*(total_officer_deaths) +
                    state_cases*(total_officer_deaths + state_deaths), 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
(BF.sig2 <- bayesfactor_parameters(sig2, null = 0))
(loo.sig2 <- loo(sig2, save_psis = T))
loo.best
```
```{r optimization 3}
loo.best <- loo.sig2
sig3 <- stan_glm(data = sys.scaled.data, 
                        formula = 
                    total_inmate_deaths~ inmate_tests*(total_officer_deaths) +
                    total_inmate_cases*(total_officer_deaths) +
                    latest_inmate_population*(total_officer_deaths + state_deaths) +
                    total_officer_cases*(total_officer_deaths) +
                    state_cases*(total_officer_deaths + state_deaths), 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
(BF.sig3 <- bayesfactor_parameters(sig3, null = 0))
(loo.sig3 <- loo(sig3, save_psis = T))
loo.best
```
```{r optimization 4}
loo.best <- loo.sig3
sig4 <- stan_glm(data = sys.scaled.data, 
                        formula = 
                    total_inmate_deaths~ inmate_tests*(total_officer_deaths) +
                    latest_inmate_population*(total_officer_deaths + state_deaths) +
                    total_officer_cases*(total_officer_deaths) +
                    state_cases*(total_officer_deaths + state_deaths), 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
(BF.sig4 <- bayesfactor_parameters(sig4, null = 0))
(loo.sig4 <- loo(sig4, save_psis = T))
loo.best
```
```{r optimization 5}
loo.best <- loo.sig4
sig5 <- stan_glm(data = sys.scaled.data, 
                        formula = 
                    total_inmate_deaths~ inmate_tests*(total_officer_deaths) +
                    latest_inmate_population*(state_deaths) +
                    total_officer_cases*(total_officer_deaths) +
                    state_cases*(total_officer_deaths + state_deaths), 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
(BF.sig5 <- bayesfactor_parameters(sig5, null = 0))
(loo.sig5 <- loo(sig5, save_psis = T))
loo.best
```
```{r optimization 6}
loo.best <- loo.sig5
sig6 <- stan_glm(data = sys.scaled.data, 
                        formula = 
                    total_inmate_deaths~ inmate_tests*(total_officer_deaths) +
                    latest_inmate_population*(state_deaths) +
                    state_cases*(total_officer_deaths + state_deaths), 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
(BF.sig6 <- bayesfactor_parameters(sig6, null = 0))
(loo.sig6 <- loo(sig6, save_psis = T))
loo.best
```
```{r optimization 7}
loo.best <- loo.sig6
sig7 <- stan_glm(data = sys.scaled.data, 
                        formula = 
                    total_inmate_deaths~ inmate_tests*(total_officer_deaths) +
                    latest_inmate_population*(state_deaths) +
                    state_cases*(state_deaths) + total_officer_deaths, 
                         family= gaussian(),
                         refresh = 0, refresh = 0, seed = 5)
(BF.sig7 <- bayesfactor_parameters(sig7, null = 0))
(loo.sig7 <- loo(sig7, save_psis = T))
loo.best
```
```{r final model}
final.model <- sig7
mcmc_areas(final.model, prob=0.90)

color_scheme_set("red")
ppc_dens_overlay(y = as.vector(final.model$y),
                 yrep = posterior_predict(final.model, draws = 50))
```

```{r}
int.1 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ inmate_tests*., 
                         family= neg_binomial_2(link="log"),
                         refresh = 0, refresh = 0, seed = 4)
round(posterior_interval(int.1, prob = 0.90), 3)
```

```{r interactions w/ total inmate cases}
int.2 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ total_inmate_cases*., 
                         family= neg_binomial_2(link="log"),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(int.2, prob = 0.90), 3)
```
```{r interactions w/ inmate population}
int.3 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ latest_inmate_population*., 
                         family= neg_binomial_2(link="log"),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(int.3, prob = 0.90), 3)
```
```{r interactions w/ officer cases}
int.4 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ total_officer_cases*., 
                         family= neg_binomial_2(link="log"),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(int.4, prob = 0.90), 3)
```
```{r interactions w/ officer deaths}
int.5 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ total_officer_deaths*., 
                         family= neg_binomial_2(link="log"),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(int.5, prob = 0.90), 3)
```
```{r interactions w/ statewide cases}
int.6 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ state_cases*., 
                         family= neg_binomial_2(link="log"),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(int.6, prob = 0.90), 3)
```
```{r interactions w/ statewide deaths}
int.7 <- stan_glm(data = nrv.scaled.data, 
                        formula = total_inmate_deaths~ state_deaths*., 
                         family= neg_binomial_2(link="log"),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(int.7, prob = 0.90), 3)
```
```{r only significant predictors}
sig.0 <- stan_glm(data = nrv.scaled.data, 
                  formula = total_inmate_deaths~ total_inmate_cases +
                    total_officer_deaths + total_officer_cases*(state_deaths)
                  + latest_inmate_population*(state_deaths) + inmate_tests, 
                         family= neg_binomial_2(link="log"),
                         refresh = 0, refresh = 0, seed = 5)
round(posterior_interval(sig.0, prob = 0.90), 3)
```
```{r}
mcmc_areas(sig.0, prob=0.90)

color_scheme_set("red")
ppc_dens_overlay(y = as.vector(sig.0$y),
                 yrep = posterior_predict(sig.0, draws = 50))
```
```{r}
(loo.sig0 <- loo(sig.0, save_psis = T))
loo.nb1
loo_compare(loo.sig0, loo.nb1)
```
```{r}
(BF.sig0 <- bayesfactor_parameters(sig.0, null = 0))
```

# Compare Loo of Fitted Normal Model v. Fitted Negative Binomial Model
```{r comparing final loo for Normal v. NB Models}
loo.sig0
loo.full.int
loo_compare(loo.sig0, loo.full.int)
```
These 2 models are not significantly different in prediction rate based on comparisons between their looic and the difference in expected log posterior density.

# Comparing Dataset after Removing Outliers 
# Compare Differences in PSIS plots for loo.sig0 & loo.full.int